---
title: "Task1 - Introduction to R Markdown"
author: "Monika Stanislovenaite, EKO"
date: "2016-02-09"
output: html_document
---
---

R Markdown is an authoring format that enables easy creation of dynamic documents, presentations, and reports from R. It combines the core syntax of markdown (an easy-to-write plain text format) with embedded R code chunks that are run so their output can be included in the final document. R Markdown documents are fully reproducible (they can be automatically regenerated whenever underlying R code or data changes).

Use an asterisk mark, to provide emphasis such as *italics* and **bold**

Create lists with a dash:

- Item 1
- Item 2
- Item 3

You can write `in-line` code with a back-tick.

```
Code blocks display
with fixed-width font
```

> Blockquotes are offset

With R Markdown, you can insert R code chunks:

```{r}
#quick summary
summary(cars)
```

You can also include plots:

```{r, echo=FALSE}
plot(cars)
```

> Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Source: http://rmarkdown.rstudio.com/
