---
title: "Task6"
author: "Monika Stanislovėnaitė, EKO"
date: "2016-04-05"
output: html_document
---
---
[@Hyndman2014a, ch. 1. Introduction to forecasting, Lab Session 1]

###Lab Session 1, uzduotis nr. 1  
Use the Dow Jones index (data set dowjones) to do the following:  
(a) Produce a time plot of the series.  
(b) Produce forecasts using the drift method and plot them.  
(c) Show that the graphed forecasts are identical to extending the line drawn between the first and last observations.  
(d) Try some of the other benchmark functions to forecast the same data set. Which do you think is best? Why?  
  
Uzduociai reikes paketo **fpp**. Duomenys - *dowjones*:
```{r}
library(fpp)
dowjones
```
  
**(a)** Produce a time plot of the series:
```{r}
plot(dowjones, col=4, lwd=2, main="Duomenu grafikas")
```
  
**(b)** Produce forecasts using the drift method and plot them. (Drift method forecasts equal to last value plus average change):
```{r}
#forecasts:
f_drift <- rwf(dowjones, drift=TRUE, h=20)  # h - number of periods for forecasting
f_drift
#plot:
plot(f_drift, col=2, lwd=2)
```
  
**(c)** Show that the graphed forecasts are identical to extending the line drawn between the first and last observations.
```{r}
#randame pirmo ir pakutinio stebejimo reiksmes:
st1 <- dowjones[1]
st78 <- dowjones[78]
# f-ja segments isbres linija tarp dvieju tasku
plot(f_drift, col=2, lwd=2)
segments(1, st1, 78, st78, col="violet")
#matome, kad is tiesu drift metodas duoda tiese, gaunama pratesiant linija, einancia per pirmo ir paskutinio stebejimo taskus.
```
  
**(d)** Try some of the other benchmark functions to forecast the same data set. Which do you think is best? Why?
```{r}
#isbandysime visus metodus - drift, mean, naive ir seasonal naive.
par(mfrow=c(2,2))
#drift metodas (forecasts equal to last value plus average change):
plot(rwf(dowjones, drift=TRUE, h=20))
#mean metodas (forecast of all future values is equal to mean of historical data):
plot(meanf(dowjones, h=20), col=5, lwd=2)
#naive metodas (forecasts equal to last observed value):
plot(naive(dowjones, h=20))
#seasonal naive metodas (forecasts equal to last value from same season):
plot(snaive(dowjones, h=20))
```
  
**Mano nuomone, prognozems siuo atveju labiausiai tinka naive metodas.** Drift metodas siulo gana staigu kilima, o matome, kad duomenyse yra buve ir mazu reiksmiu. Mean metodas duoda staigu kritima. Seasonal naive metodas cia netinkamas, nes duomenyse nematome sezoniskumo.

